---
title: "Exam 2"
author: 'Ziyuan Shen'
date: "11/14/2019"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
```

```{r library}
library(httr)
library(tidyverse)
library(jsonlite)
library(shiny)
```

API key is stored in a json file.
```{r get-api-key}
# read in api key
news_key <- fromJSON("api_keys.json")$news_key
```

## Task 1
#### Task 1.1
Get general and business sources respectively.
```{r task1-1-get-data}
url_general <- paste0("https://newsapi.org/v2/sources?category=general&language=en&country=us&apiKey=", news_key)
url_business <- paste0("https://newsapi.org/v2/sources?category=business&language=en&country=us&apiKey=", news_key)

r_general <- content(GET(url_general))
r_business <- content(GET(url_business))
```

Combine the sources and get a cleaned dataframe that lists ids.
```{r task1-1-parse-data}
df_general <- as.tibble(do.call(rbind, r_general$sources))
df_business <- as.tibble(do.call(rbind, r_business$sources))
df_sources <- rbind(df_general, df_business)
df_sources <- df_sources %>%
  unnest(colnames(df_sources)) %>%
  filter(name %in% c('CNN', 'Fox News', 'The Wall Street Journal', 'Reuters')) %>%
  select(name, id)
df_sources
```

#### Task 1.2
Get headlines data that contains taxes.
```{r task1-2-get-data}
url_taxes <- paste0("https://newsapi.org/v2/top-headlines?country=us&q=taxes&apiKey=", news_key)
r_taxes <- content(GET(url_taxes))
```

Parse the data into cleaned dataframe. Return 'No headline found!' if dataframe is empty.
```{r task1-2-parse-data}
df_taxes <- as.tibble(do.call(rbind, r_taxes$articles))
df_taxes$id <- lapply(df_taxes$source, `[[`, 'id')
df_taxes$name <- lapply(df_taxes$source, `[[`, 'name')
if (nrow(df_taxes)==0) {
  print('No headline found!')
} else {
  df_taxes <- df_taxes %>%
    select(-source) %>%
    unnest() %>%
    filter(name %in% c('CNN', 'Fox News', 'The Wall Street Journal', 'Reuters')) %>%
    select(name, title)
  print(df_taxes)
}
```

#### Task 1.3
Get all headlins that contain healthcare in title.
```{r task1-3-get-data}
url_healthcare <- paste0("https://newsapi.org/v2/everything?qInTitle=healthcare&sources=cnn,fox-news,the-wall-street-journal,reuters&from=2019-11-01&to=2019-11-01&language=en&apiKey=", news_key)
r_healthcare <- content(GET(url_healthcare))
```

Parse the data into cleaned dataframe. Return 'No headline found!' if dataframe is empty.
```{r task1-3-parse-data}
df_healthcare <- as.tibble(do.call(rbind, r_healthcare$articles))
df_healthcare$id <- lapply(df_healthcare$source, `[[`, 'id')
df_healthcare$name <- lapply(df_healthcare$source, `[[`, 'name')
if (nrow(df_healthcare)==0) {
  print('No headline found!')
} else {
  df_healthcare <- df_healthcare %>%
    select(-source) %>%
    unnest() %>%
    select(name, title)
  print(df_healthcare)
}
```


## Task 2
Helper functions are created for each API endpoint. Inputs are checked for both type correctness and value correctness. If inputs are incorrect, function will be stopped with a message indicating the reason.
```{r task2-1}
get_sources <- function(category='', api_key=NULL) {
  if (!is.character(category) | length(category)==0) {
    stop('Please input valid category')
  }
  if (length(category)>1) {
    stop('Only one category is accepted')
  }
  if (!category %in% c('business', 'entertainment', 'general', 'health', 'science', 'sports', 'technology')) {
    stop('category must be one of business, entertainment, general, health, science, sports, technology')
  }
  if (is.null(api_key)) {
    stop('api_key must not be NULL')
  }
  url <- paste0("https://newsapi.org/v2/sources?category=", category, "&language=en&country=us&apiKey=", api_key)
  r <- content(GET(url))
  if (r$status == 'error') {
    stop(paste0(r$code, '\n', 'Error Message: ', r$message))
  } else {
    df <- as.tibble(do.call(rbind, r$sources))
    if (nrow(df) == 0) {
      return(df)
    } else {
      df <- df %>%
        unnest(colnames(df)) %>%
        filter(name %in% c('CNN', 'Fox News', 'The Wall Street Journal', 'Reuters'))
      return(df)
    }
  }
}
```

```{r task2-2}
check_sources <- function(sources) {
  if (!is.character(sources) | length(sources)==0) {
    stop('please input valid sources')
  }
  for (i in seq_along(sources)) {
    if (!sources[i] %in% c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters')) {
      stop('sources should only be a subset of cnn, fox-news, the-wall-street-journal, and reuters')
    }
  }
}

get_headlines <- function(sources=c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters'), q="", page_size=20, page=1, api_key=NULL) {
  check_sources(sources)
  #if (!is.character(q)) {
    #stop('q should be a string')
  #}
  if (!is.numeric(page_size) | page_size<=0 | page_size>100 | as.integer(page_size)!=page_size) {
    stop('page_size should be a positive integer less or equal than 100')
  }
  if (!is.numeric(page) | page<=0 | as.integer(page)!=page) {
    stop('page should be a positive integer')
  }
  if (is.null(api_key)) {
    stop('api_key must not be NULL')
  }
  url <- paste0("https://newsapi.org/v2/top-headlines?country=us&q=", q, "&pageSize=", page_size, "&page=", page, "&apiKey=", api_key)
  r <- content(GET(url))
  if (r$status == 'error') {
    stop(paste0(r$code, '\n', 'Error Message: ', r$message))
  } else {
    df <- as.tibble(do.call(rbind, r$articles))
    if (nrow(df) == 0) {
      return(df)
    } else {
      df$id <- lapply(df$source, `[[`, 'id')
      df$name <- lapply(df$source, `[[`, 'name')
      df <- df %>%
        select(-source)
      df <- df %>%
        unnest(colnames(df)) %>%
        filter(id %in% sources)
      return(df)
    }
  }
}
```


```{r task2-3}
# additional helper function that checks validity of date time input (time is optional as required by the api endpoint)
get_dttm <- function(dttm) {
  if ((!is.null(dttm)) & (dttm!="")) {
    if (!is.character(dttm)) {
      stop('from and to should be a valid date or date time if specificed')
    } else if (!is.na(as.POSIXct(dttm, format="%Y-%m-%dT%H:%M:%OS"))) {
      return(format(as.POSIXct(dttm, format="%Y-%m-%dT%H:%M:%OS"), "%Y-%m-%dT%H:%M:%OS"))
    } else if (!is.na(as.Date(dttm, format="%Y-%m-%d"))) {
      return(as.Date(dttm, format="%Y-%m-%d"))
    } else {
      stop('from and to should be a valid date or date time in the form like 2019-11-18 or 2019-11-18T07:13:34')
    }
  } else {
    return(NULL)
  }
}

get_historic <- function(q="", q_title="", sources=c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters'), from=NULL, to=NULL, sort_by="publishedAt", page_size=20, page=1, api_key=NULL) {
  check_sources(sources)
  sources <- paste(sources, collapse=',')
  from <- get_dttm(from)
  to <- get_dttm(to)
  if (!sort_by %in% c('relevancy', 'popularity', 'publishedAt')) {
    stop('sort_by must be one of relevancy, popularity, publishedAt')
  }
  if (!is.numeric(page_size) | page_size<=0 | page_size>100 | as.integer(page_size)!=page_size) {
    stop('page_size should be a positive integer less or equal than 100')
  }
  if (!is.numeric(page) | page<=0 | as.integer(page)!=page) {
    stop('page should be a positive integer')
  }
  if (is.null(api_key)) {
    stop('api_key must not be NULL')
  }
  url <- paste0("https://newsapi.org/v2/everything?q=", q, "&qInTitle=", q_title, "&sources=", sources, "&from=", from, "&to=", to, "&language=en&sortBy=", sort_by, "&pageSize=", page_size, "&page=", page, "&apiKey=", api_key)
  r <- content(GET(url))
  if (r$status == 'error') {
    stop(paste0(r$code, '\n', 'Error Message: ', r$message))
  } else {
    df <- as.tibble(do.call(rbind, r$articles))
    if (nrow(df)==0) {
      return(df)
    } else {
      df$id <- lapply(df$source, `[[`, 'id')
      df$name <- lapply(df$source, `[[`, 'name')
      df <- df %>%
        select(-source)
      df <- df %>%
        unnest(colnames(df)) 
      return(df)
    }
  }
}
```



## Task 3
Create a News Hub shiny app. Navigation bar page is employed. Three helper functions from Task 2 are used and all inputs aruguments can be specified by user. API data for the three functions is retrived and updated only when the update button in each tabPanel is clicked. For get_sources, a dataframe table is displayed. For get_headlines and get_historic, each headline is displayed with the image, title and description. Also, each headline has a Read More button. When the button is clicked, more information (author, content, link) will pop up in a modal dialog box. Conditaionl panel and reactive outputs are employed as the number of headlines retrieved each time is different.
```{r task3}
shinyApp (
  ui <- fluidPage(
    navbarPage(title="News Hub",
      tabPanel("Get Sources",
          fluidRow(
            column(width=2,
              radioButtons("sources_category", "Category", c('business', 'entertainment', 'general', 'health', 'science', 'sports', 'technology'))
            ),
            column(width=1,
              br(),br(),br(),br(),br(),br(),br(),br(),
              actionButton("sources_update", "Update")
            )
          ),
          fluidRow(
            br(),br(),br()
          ),
          fluidRow(
            dataTableOutput("sources")
          )
      ),
      tabPanel("Get Headlines",
        sidebarLayout(
          sidebarPanel(
            checkboxGroupInput("headlines_sources", "Sources", c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters'), selected=c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters')),
            textInput("headlines_q", "Search All", ""),
            sliderInput("headlines_pagesize", "pagesize", 20, 100, 20),
            sliderInput("headlines_page", "page", 1, 20, 1),
            actionButton("headlines_update", "Update")
          ),
          mainPanel(
            lapply(1:100, function(i) {
              conditionalPanel(condition = paste0("output.headlines_nrow>=", i),
                fluidRow(
                  fluidRow(
                    column(width=2,
                      htmlOutput(paste0("headlines_row", i, "_image"))
                    ),
                    column(width=10,
                      htmlOutput(paste0("headlines_row", i)),
                      br(),
                      actionButton(paste0("headlines_row", i, "_readmore"), "Read More")
                    )
                  ),
                  fluidRow(
                    br()
                  )
                )
              )
            }
            )
          )
        )
      ),
      tabPanel("Get Historic",
        sidebarLayout(
          sidebarPanel(
            textInput("historic_q", "Search All", ""),
            textInput("historic_qtitle", "Search In Title", ""),
            checkboxGroupInput("historic_sources", "Sources", c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters'), selected=c('cnn', 'fox-news', 'the-wall-street-journal', 'reuters')),
            textInput("historic_from", "From", placeholder="2019-11-18T07:13:34"),
            textInput("historic_to", "To", placeholder="2019-11-18T07:13:34"),
            radioButtons("historic_sortby", "Sort By", c('publishedAt', 'relevancy', 'popularity')),
            sliderInput("historic_pagesize", "pagesize", 20, 100, 20),
            sliderInput("historic_page", "page", 1, 20, 1),
            actionButton("historic_update", "Update")
          ),
          mainPanel(
            lapply(1:100, function(i) {
              conditionalPanel(condition = paste0("output.historic_nrow>=", i),
                fluidRow(
                  fluidRow(
                    column(width=2,
                      htmlOutput(paste0("historic_row", i, "_image"))
                    ),
                    column(width=10,
                      htmlOutput(paste0("historic_row", i)),
                      br(),
                      actionButton(paste0("historic_row", i, "_readmore"), "Read More")
                    )
                  ),
                  fluidRow(
                    br()
                  )
                )
              )
            }
            )
          )
        )
      )
    )
  ),
  
  server <- function(input, output) {
    output$sources <- renderDataTable({
      input$sources_update
      sources_output <- isolate(get_sources(category=input$sources_category, api_key=news_key))
      sources_output
    })
    
    headlines_output <- reactive({
      input$headlines_update
      isolate(get_headlines(sources=input$headlines_sources, q=input$headlines_q, page_size=input$headlines_pagesize, page=input$headlines_page, api_key=news_key))
    })

    output$headlines_nrow <- reactive({
      nrow(headlines_output())
    })
    
    outputOptions(output, "headlines_nrow", suspendWhenHidden = FALSE) 
    
    lapply(1:100, function(i) {
        output[[paste0("headlines_row", i, "_image")]] <- renderText({
          paste0('<img src="',headlines_output()$urlToImage[i], '" style="width:100%;height:60%;"', '>')
        })
        output[[paste0("headlines_row", i)]] <- renderText({
          paste0(
            paste0('<div align="left"><strong>', headlines_output()$title[i], ' </strong></div>'),
            paste0('<div align="left">', format(as.POSIXct(headlines_output()$publishedAt[i]), "%Y-%m-%d %H:%M:%OS"), '</div>'),
            paste0('<div align="left">', headlines_output()$description[i], '</div>')
          )
        })
        observeEvent(input[[paste0("headlines_row", i, "_readmore")]], {
          showModal(modalDialog(
            title = headlines_output()$title[i],
            HTML(paste0('Author: ', headlines_output()$author[i], '<br><br> Content: <br>', headlines_output()$content[i], '<br><br>Link:<br>', headlines_output()$url[i]))
          ))
        })
      }
    )
    
    historic_output <- reactive({
      input$historic_update
      isolate(get_historic(q=input$historic_q, q_title=input$historic_qtitle, sources=input$historic_sources, from=input$historic_from, to=input$historic_to, sort_by=input$historic_sortby, page_size=input$historic_pagesize, page=input$historic_page, api_key=news_key))
    })

    output$historic_nrow <- reactive({
      nrow(historic_output())
    })
    
    outputOptions(output, "historic_nrow", suspendWhenHidden = FALSE) 
    
    lapply(1:100, function(i) {
        output[[paste0("historic_row", i, "_image")]] <- renderText({
          paste0('<img src="',historic_output()$urlToImage[i], '" style="width:100%;height:60%;"', '>')
        })
        output[[paste0("historic_row", i)]] <- renderText({
          paste0(
            paste0('<div align="left"><strong>', historic_output()$title[i], ' </strong></div>'),
            paste0('<div align="left">', format(as.POSIXct(historic_output()$publishedAt[i]), "%Y-%m-%d %H:%M:%OS"), '</div>'),
            paste0('<div align="left">', historic_output()$description[i], '</div>')
          )
        })
        observeEvent(input[[paste0("historic_row", i, "_readmore")]], {
          showModal(modalDialog(
            title = historic_output()$title[i],
            HTML(paste0('Author: ', historic_output()$author[i], '<br><br> Content: <br>', historic_output()$content[i], '<br><br>Link:<br>', historic_output()$url[i]))
          ))
        })
      }
    )

  }
)
```


